BOLD Parameter Estimation Using Particle Filters

### 2010-08-09 doc/thesis/conclusion.tex
added a signle section title and recompiled
updated conclusion

Although point estimates of parameters are not dependable estimates of the 
true parameters, they are by no means usefulness. Analysis of differences in parameters
could be medically relevant. Additionally, a single estimate for parameters is still able
to form a good estimate of the BOLD signal. This factor adds bridge to 
earlier types of activation tests, such as the statistical
parametric mapping. As the results in \autoref{sec:RealData} show, the heatmaps, especially
those of mutual information closely resembled the results of SPM. While the activation
tests were more sensitive, they were also more prone to false positive. However, it is
worth noting that certain enhancements could be useful in reducing false positives; for instance
by using the maximum likelihood of the final distribution, or even the median. The 
usefulness of these techniques depend on the degree to which the algorithm converged, 
and thus on the experimental design. Future work may shed further light on these
techniques.

% Pros/Cons
%% Cons
%%% Computation longer than SPM
%%% Harder to interpret
%% Pro
%%% Non-parametric
%%% Real time
%%% Full Posterior
%%% More intuitive
\section{Particle Filter Approach}
The Particle Filter algorithm was originally designed for on-line parameter 
estimation. For this reason, there is no guarantee of optimality or even 
convergence for finite measurements. However, for the BOLD nonlinear ODE there
can be no guarantee of an optimal solution. A guaranteed local minimum,
which nonlinear least squares provides, however would be helpful. One difficulty
with the use of a particle filter with finite data is the calculation of a good
weight function $P(y_k | x_k)$  that will converge at a decent rate. If the weight
function does not sufficiently differentiate particles, the final distribution will 
no be significantly different from the prior distribution. On the other hand, if the 
weight function is too thin, it will unfairly eliminate viable particles. Given sufficient
measurements it is better to let the algorithm take longer to converge, because the
convergence will be better (more accurate). The particle filter takes longer
to run than Volterra approximation method from \autoref{sec:Background Linear Approximation};
however, it is free from the uncertainty of whether a quadratic approximation is 
sufficient for the BOLD model.

@@

A more practical concern with the particle filter is that it is mathematically
simple. Only a basic understanding of Bayesian statistics is needed to understand
how the particle filter works. This is in contrast to the Volterra based approach
of \cite{Friston2000}, which is quite complex. Additionally very few assumptions
are needed for the particle filter. It is a common problem in traditional statistics 
for assumptions to be unrealistic which means the results may be invalid or at 
least in question. 
Fewer and more realistic assumptions mean that the particle filter is more robust
to unforeseen difficulties in FMRI data.

@@

% Future works
%% Extensive test of quality of volterra kernel estimate from friston
%% Better de-trending
%% Automatic estimation of measurement error.
%todo

As result, it is possible to use
the particle filter to localize areas where a known stimuli most directly drives
neural activation. In the future it will be possible to use the estimated states
($v,q,s,f$) to drive other models and learn more about how regions of the brain
interact.

>>>>

Although point estimates of parameters are not dependable estimates of the 
true parameters, they are by no means useless. Analysis of differences in parameters
could be medically relevant. Additionally, a single estimate for parameters is still able
to form a good estimate of the BOLD signal. This factor forms a bridge to 
earlier types of activation tests, such as the statistical
parametric mapping. As the results in \autoref{sec:RealData} show, the heatmaps, especially
those of mutual information closely resembled the results of SPM. While the activation
tests were more sensitive, they were also more prone to false positive. However, it is
worth noting that certain enhancements could be useful in reducing false positives; for instance
by using the maximum likelihood of the final distribution, or even the median. The 
usefulness of these techniques depend on the degree to which the algorithm converged, 
and thus on the experimental design. Future work may shed further light on these
techniques.

% Pros/Cons
%% Cons
%%% Computation longer than SPM
%%% Harder to interpret
%% Pro
%%% Non-parametric
%%% Real time
%%% Full Posterior
%%% More intuitive
\section{Review of Particle Filter Approach}
The Particle Filter algorithm was originally designed for on-line parameter 
estimation. For this reason, there is no guarantee of optimality or even 
convergence for finite measurements. However, for the BOLD nonlinear ODE 
this is less of a concern than it might first appear to be. For this
particular problem there can be no guarantee of a global minimum, and although
other techniques guarantee a local minimum, the particle filter doesn't settle 
to one of these local minima because it can settle into multiple ones. This
difference typifies the major difference between the particle filter and 
competing approaches.

One difficulty
with the use of a particle filter with finite data is the calculation of a good
weight function $P(y_k | x_k)$  that will converge at a decent rate. If the weight
function does not sufficiently differentiate particles, the final distribution will 
no be significantly different from the prior distribution. On the other hand, if the 
weight function is too thin, it will unfairly eliminate viable particles. Given sufficient
measurements it is better to let the algorithm take longer to converge, because the
convergence will be better (more accurate). The particle filter takes longer
to run than Volterra approximation method from \autoref{sec:Background Linear Approximation};
however, it is free from the uncertainty of whether a quadratic approximation is 
sufficient for the BOLD model.

@@

A more practical benefit with the particle filter is that it is mathematically
simple. An understanding of Bayesian statistics is all that is necessary to understand
how the particle filter works. This is in contrast to the Volterra based approach
of \cite{Friston2000}, which is quite complex. Additionally very few assumptions
are needed for the particle filter. It is a common problem in traditional statistics 
for assumptions to be unrealistic which means the results may be invalid or at 
least in question. 
Fewer and more realistic assumptions mean that the particle filter is more robust
to unforeseen difficulties in FMRI data.

@@

% Future works
%% Extensive test of quality of volterra kernel estimate from friston
%% Better de-trending
%% Automatic estimation of measurement error.
%% Using s as the input to other regions. 
%% Extensive parameter estimation to find a correct prior
%% Increased constraints on parameters using physiological data
%% Incorporating Blood Flow/Volume data as separate measurement
\section{Future Works}
